Capstone Movielens Report

### 2025-05-09 reports/capstone-movielens-report.draft4.Rmd
Report updated: UME Model Regularization feature.

  1. **Preconfiguration:** Preliminary determination of the optimal range of `lambda` values for the `5-Fold Cross Validation` samples;
  2. **Fine-tuning:** figuring out the value of `lambda` that minimizes the model's RMSE.

@@

```{r eval=FALSE}
}

@@

This function calculates the _RMSE_ for the given $\lambda$ parameter value:
```{r eval=FALSE}
regularize.test_lambda.UM_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UM_effect.cv
`lambda` is `NA`")
  }
  um_effect <- train_user_movie_effect.cv(lambda)
  calc_user_movie_effect_RMSE.cv(um_effect)
}
```

>>>>

  1. **Preconfiguration:** Preliminary determination of the optimal range of $\lambda$ values for the `5-Fold Cross Validation` samples;
  2. **Fine-tuning:** figuring out the value of $\lambda$ that minimizes the model's RMSE.

@@

```{r eval=FALSE}
tune.model_param <- function(param_values, 
                             fn_tune.test.param_value, 
                             break.if_min = TRUE,
                             steps.beyond_min = 2){
  n <- length(param_values)
  param_vals_tmp <- numeric()
  RMSEs_tmp <- numeric()
  RMSE_min <- Inf
  i_max.beyond_RMSE_min <- Inf
  prm_val.best <- NA

  put_log("Function: `tune.model_param`:
param_values:")
  put(param_values)

  for (i in 1:n) {
    put_log1("Function: `tune.model_param`:
Iteration %1", i)
    prm_val <- param_values[i]
    put_log1("Function: `tune.model_param`:
prm_val: %1", prm_val)
    param_vals_tmp[i] <- prm_val

    put_log2("Function: `tune.model_param`:
param_vals_tmp[%1]: %2", i, param_vals_tmp[i])
    put_log1("Function: `tune.model_param`:
param_vals_tmp length: %1", length(param_vals_tmp))
    put(param_vals_tmp)

    RMSE_tmp <- fn_tune.test.param_value(prm_val)

    put_log1("Function: `tune.model_param`:
RMSE_tmp: %1", RMSE_tmp)
    RMSEs_tmp[i] <- RMSE_tmp

    put_log2("Function: `tune.model_param`:
RMSEs_tmp[%1]: %2", i, RMSEs_tmp[i])
    put_log1("Function: `tune.model_param`:
RMSEs_tmp length: %1", length(RMSEs_tmp))
    put(RMSEs_tmp)

    plot(param_vals_tmp[RMSEs_tmp > 0], RMSEs_tmp[RMSEs_tmp > 0])
    # browser()

    if(RMSE_tmp > RMSE_min){
      warning("Function: `tune.model_param`:
`RSME` reached its minimum: ", RMSE_min, "
for parameter value: ", prm_val)
      put_log2("Function: `tune.model_param`:
Current `RMSE` value is %1 related to parameter value: %2",
               RMSE_tmp,
               prm_val)

      if (i > i_max.beyond_RMSE_min) {
        warning("Function: `tune.model_param`:
Operation is breaked (after `RSME` reached its minimum) on the following step: ", i)
        # browser()
        break
      }
      # browser()
      next
    }

    RMSE_min <- RMSE_tmp
    prm_val.best <- prm_val

    if (break.if_min) {
      i_max.beyond_RMSE_min <- i + steps.beyond_min
    }
    # browser()
  }

  param_values.best_result <- c(param.best_value = prm_val.best, 
                                best_RMSE = RMSE_min)

  put_log1("Function: `tune.model_param`:
Completed with RMSEs_tmp length: %1", length(RMSEs_tmp))
  list(tuned.result = data.frame(RMSE = RMSEs_tmp,
                                parameter.value = param_vals_tmp),
       best_result = param_values.best_result)
}

@@

This function calculates _RMSE_ of the _UME Model_ using _5-Fold CV_ method for the given $\lambda$ parameter value:
```{r eval=FALSE}
regularize.test_lambda.UM_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UM_effect.cv
`lambda` is `NA`")
  }
  um_effect <- train_user_movie_effect.cv(lambda)
  calc_user_movie_effect_RMSE.cv(um_effect)
}
```

### 2025-05-09 reports/capstone-movielens-report.draft4.tex
Report updated: UME Model Regularization feature.

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-12-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-39-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-44-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-47-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-52-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-54-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-65-1.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Preconfiguration:} Preliminary determination of the optimal
  range of \texttt{lambda} values for the
  \texttt{5-Fold\ Cross\ Validation} samples;
\item
  \textbf{Fine-tuning:} figuring out the value of \texttt{lambda} that
  minimizes the model's RMSE.
\end{enumerate}

@@

\begin{Shaded}
\begin{Highlighting}[]
\ErrorTok{\}}
\end{Highlighting}
\end{Shaded}

@@

This function calculates the \emph{RMSE} for the given \(\lambda\)
parameter value:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-85-1.pdf}

>>>>

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-11-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-38-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-43-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-46-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-51-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-53-1.pdf}

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-64-1.pdf}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Preconfiguration:} Preliminary determination of the optimal
  range of \(\lambda\) values for the \texttt{5-Fold\ Cross\ Validation}
  samples;
\item
  \textbf{Fine-tuning:} figuring out the value of \(\lambda\) that
  minimizes the model's RMSE.
\end{enumerate}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{tune.model\_param }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(param\_values, }
\NormalTok{                             fn\_tune.test.param\_value, }
                             \AttributeTok{break.if\_min =} \ConstantTok{TRUE}\NormalTok{,}
                             \AttributeTok{steps.beyond\_min =} \DecValTok{2}\NormalTok{)\{}
\NormalTok{  n }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(param\_values)}
\NormalTok{  param\_vals\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{()}
\NormalTok{  RMSEs\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{numeric}\NormalTok{()}
\NormalTok{  RMSE\_min }\OtherTok{\textless{}{-}} \ConstantTok{Inf}
\NormalTok{  i\_max.beyond\_RMSE\_min }\OtherTok{\textless{}{-}} \ConstantTok{Inf}
\NormalTok{  prm\_val.best }\OtherTok{\textless{}{-}} \ConstantTok{NA}

  \FunctionTok{put\_log}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_values:"}\NormalTok{)}
  \FunctionTok{put}\NormalTok{(param\_values)}

  \ControlFlowTok{for}\NormalTok{ (i }\ControlFlowTok{in} \DecValTok{1}\SpecialCharTok{:}\NormalTok{n) \{}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Iteration \%1"}\NormalTok{, i)}
\NormalTok{    prm\_val }\OtherTok{\textless{}{-}}\NormalTok{ param\_values[i]}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{prm\_val: \%1"}\NormalTok{, prm\_val)}
\NormalTok{    param\_vals\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ prm\_val}

    \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_vals\_tmp[\%1]: \%2"}\NormalTok{, i, param\_vals\_tmp[i])}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{param\_vals\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(param\_vals\_tmp))}
    \FunctionTok{put}\NormalTok{(param\_vals\_tmp)}

\NormalTok{    RMSE\_tmp }\OtherTok{\textless{}{-}} \FunctionTok{fn\_tune.test.param\_value}\NormalTok{(prm\_val)}

    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSE\_tmp: \%1"}\NormalTok{, RMSE\_tmp)}
\NormalTok{    RMSEs\_tmp[i] }\OtherTok{\textless{}{-}}\NormalTok{ RMSE\_tmp}

    \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSEs\_tmp[\%1]: \%2"}\NormalTok{, i, RMSEs\_tmp[i])}
    \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{RMSEs\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(RMSEs\_tmp))}
    \FunctionTok{put}\NormalTok{(RMSEs\_tmp)}

    \FunctionTok{plot}\NormalTok{(param\_vals\_tmp[RMSEs\_tmp }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{], RMSEs\_tmp[RMSEs\_tmp }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{])}
    \CommentTok{\# browser()}

    \ControlFlowTok{if}\NormalTok{(RMSE\_tmp }\SpecialCharTok{\textgreater{}}\NormalTok{ RMSE\_min)\{}
      \FunctionTok{warning}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{\textasciigrave{}RSME\textasciigrave{} reached its minimum: "}\NormalTok{, RMSE\_min, }\StringTok{"}
\StringTok{for parameter value: "}\NormalTok{, prm\_val)}
      \FunctionTok{put\_log2}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Current \textasciigrave{}RMSE\textasciigrave{} value is \%1 related to parameter value: \%2"}\NormalTok{,}
\NormalTok{               RMSE\_tmp,}
\NormalTok{               prm\_val)}

      \ControlFlowTok{if}\NormalTok{ (i }\SpecialCharTok{\textgreater{}}\NormalTok{ i\_max.beyond\_RMSE\_min) \{}
        \FunctionTok{warning}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Operation is breaked (after \textasciigrave{}RSME\textasciigrave{} reached its minimum) on the following step: "}\NormalTok{, i)}
        \CommentTok{\# browser()}
        \ControlFlowTok{break}
\NormalTok{      \}}
      \CommentTok{\# browser()}
      \ControlFlowTok{next}
\NormalTok{    \}}

\NormalTok{    RMSE\_min }\OtherTok{\textless{}{-}}\NormalTok{ RMSE\_tmp}
\NormalTok{    prm\_val.best }\OtherTok{\textless{}{-}}\NormalTok{ prm\_val}

    \ControlFlowTok{if}\NormalTok{ (break.if\_min) \{}
\NormalTok{      i\_max.beyond\_RMSE\_min }\OtherTok{\textless{}{-}}\NormalTok{ i }\SpecialCharTok{+}\NormalTok{ steps.beyond\_min}
\NormalTok{    \}}
    \CommentTok{\# browser()}
\NormalTok{  \}}

\NormalTok{  param\_values.best\_result }\OtherTok{\textless{}{-}} \FunctionTok{c}\NormalTok{(}\AttributeTok{param.best\_value =}\NormalTok{ prm\_val.best, }
                                \AttributeTok{best\_RMSE =}\NormalTok{ RMSE\_min)}

  \FunctionTok{put\_log1}\NormalTok{(}\StringTok{"Function: \textasciigrave{}tune.model\_param\textasciigrave{}:}
\StringTok{Completed with RMSEs\_tmp length: \%1"}\NormalTok{, }\FunctionTok{length}\NormalTok{(RMSEs\_tmp))}
  \FunctionTok{list}\NormalTok{(}\AttributeTok{tuned.result =} \FunctionTok{data.frame}\NormalTok{(}\AttributeTok{RMSE =}\NormalTok{ RMSEs\_tmp,}
                                \AttributeTok{parameter.value =}\NormalTok{ param\_vals\_tmp),}
       \AttributeTok{best\_result =}\NormalTok{ param\_values.best\_result)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

This function calculates \emph{RMSE} of the \emph{UME Model} using
\emph{5-Fold CV} method for the given \(\lambda\) parameter value:

@@

\includegraphics{capstone-movielens-report.draft4_files/figure-latex/unnamed-chunk-84-1.pdf}
